Think deep for reef resilience

### 2025-08-06 paper/paper_draft.qmd
Minor formatting
Separate and make explicit section on DHWs
@eco-ben Please note my comments in this commit too :)

I think we do need more detail on SSP vs RCP, and how DHWs were estimated. Details on DHWs could be addressed by pointing to the relevant analysis script.

## Environmental modelling context

The spatial extent of the GBR reefs represented in this investigation follows the reef
definitions available via the Great Barrier Reef Marine Park Authority [@greatba2022]. In
total this covers 3806 reef areas across the GBR marine park. Reefs that are outside the
GBRMPA management areas or are not associated with a GBRMPA reefal bioregion were removed
from the investigation, leaving 3531 reefs. Annually and spatially averaged Degree Heating
Week (DHW) data were used to represent future thermal stress at each reef location.

Statistically downscaled 1km resolution Sea Surface Temperature (SST) data for the GBR were
obtained from the HighResCoralStress database [@dixon2022; @dixon2023]. Maximum monthly mean
DHW levels were calculated from the SST data following NOAA methodology [@liu2017;
@skirving2020]. DHWs were represented by a selection of five Global Climate Models (GCMs)
from the Sixth Phase of the Coupled Model Intercomparison Project (CMIP6). These five GCMs
were EC-Earth3-Veg, ACCESS-CM2, ACCESS-ESM1-5, NorESM2-MM and GDFL-CM4, and their inclusion
were based on evaluation by Grose et al. [-@grose2023]. Grose et al. [-@grose2023] found
these GCMs to rank highly, among 50 GCMs, in recreation of historical atmospheric
temperature and precipitation conditions in the Australasian region. While this assessment
does not consider oceanic data, Grose et al. \[\@-grose2023\] provide a region-specific
selection mechanism which we use as an indicator of general performance of GCMs for the
region.

Equilibrium climate sensitivity (ECS) is defined, by the IPCC, as the long term warming
caused by a doubling of carbon dioxide above its pre-industrial concentration
[@forster2021]. The five GCMs span a range of ECS values within the IPCC assessed likely
range, and GCM development families, allowing for a range of possible climate futures to be
explored, capturing the uncertainty in future responses to CO2 increases (@sfig-ecs;
[@grose2023; @shepherd2018]). The following analyses were performed using each GCM
separately. SSP 2-4.5 was used as a plausible emission pathway under current emission
trajectories, with similar atmospheric CO2 levels remaining until 2050 [@hausfather2020;
IPCC -@climate2023].

@@

>>>>

## Modelling context

The spatial extent of the GBR reefs represented in this investigation follows the reef
definitions available via the Great Barrier Reef Marine Park Authority [@greatba2022]. In
total, the dataset represents 3806 reef areas across the GBR marine park. Reefs that are
outside the GBRMPA management areas or are not associated with a GBRMPA reefal bioregion
were removed from the investigation, leaving 3531 reefs. Annually and spatially averaged
Degree Heating Week (DHW) data were used to represent future thermal stress at each reef
location.

@@

## Degree Heating Week Trajectories

Future projections of maximum yearly Degree Heating Weeks (DHWs) for each reef were
estimated from Sea Surface Temperature (SST) data for the GBR. Statistically downscaled 1km
resolution SST data for five Global Climate Models (GCMs) were obtained from the
HighResCoralStress database [@dixon2022; @dixon2023]. These five GCMs were EC-Earth3-Veg,
ACCESS-CM2, ACCESS-ESM1-5, NorESM2-MM and GDFL-CM4 and are a part of the Sixth Phase of the
Coupled Model Intercomparison Project (CMIP6) based on evaluation conducted by Grose et al.
[-@grose2023]. These GCMs were found to rank highly in the recreation of historical
atmospheric temperature and precipitation conditions in the Australasian region. While this
assessment does not consider oceanic data, Grose et al. [-@grose2023] provide a
region-specific selection mechanism which we use as an indicator of general performance of
GCMs for the region.

Maximum monthly mean DHW levels were calculated from the SST data following NOAA methodology
[@liu2017; @skirving2020], from which maximum yearly DHWs were determined.
**\[TI: Could you confirm the above is described correctly?\]**

The five selected GCMs span a range of Equilibrium Climate Sensitivity (ECS) values within
the IPCC assessed likely range, and GCM development families. The IPCC defines ECS as the
long term warming caused by a doubling of carbon dioxide above its pre-industrial
concentration [@forster2021]. The five GCMs therefore allow for a range of possible climate
futures to be explored, capturing the uncertainty in future responses to CO2 increases
(@sfig-ecs; [@grose2023; @shepherd2018]).

Shared Socio-economic Pathway 2 (replacing the older RCP 4.5) was used to represent the
plausible emission pathway under current emission trajectories, with similar atmospheric CO2
levels remaining until 2050 [@hausfather2020; IPCC -@climate2023].
\[**TI note: The above needs to be expanded a little.
IPCC has not been mentioned before at this point either**\]

\[**TI: Needs a figure showing DHW trajectories for GBR!**\]

\[**TI: Needs a table clearly communicating each GCM, ECS value, and what each represent
(e.g., usual, optimistic scenario, etc)**\]
